### 2024-04-22 workspace/heyuan.Rmd
Add files via upload

```{r}
library(tidyverse)
library(tseries)
#install.packages("ggfortify")
library(ggfortify)
library(pomp)
```

#### read in the data & data description
```{r}
data_path <- '/Users/ruojunliu/Desktop/STATS 531 - Time Series/pomp_final/data/data.txt'
df <- read.table(data_path, header = TRUE, sep = "\t", na.strings = "NA", dec = ".", strip.white = TRUE)
head(df)

```

142 years (1872-2013) of willow ptarmigan hunting records in southeastern Norway

CPUE: the number of birds shot per hunter per day, as the primary metric for assessing population changes.

NAO: A positive NAO index favors mild winters in the U.S. East. Negative values favor stronger cold-air outbreaks and increased storminess in the eastern U.S.

temp.PREINC: temperature, pre-incubation, 5/1 - 6/2
temp.INC: temperature, incubation, 6/3 - 6/24
temp.BROOD: temperature, brooding & rearing, 6/25 - 7/15
temp.HATCH: temperature, mean hatching date, 7/21 - 7/27
prefix prec.: precipitation

opening data hunt(julian day): the starting day of hunting season is named as Julian Day, which is varied across 142 years. It is mainly in mid to late August up to the 1920's and 10th/15th September in later years.

# Data Analysis

#### Trend Analysis
Temporal Patterns: CPUE showed significant declines from early records to recent years, with notable shifts in population cycles and their amplitude.

Impact of Alternative Prey: The availability of rodents, which serve as alternative prey for predators that also hunt ptarmigan, was strongly linked to fluctuations in ptarmigan numbers. Years with high rodent populations typically saw higher CPUE for ptarmigan, suggesting lower predatory pressure on them.

#### Correlation
For the ACF plot, the initial 40 lags show a strong correlation with past values, diminishing quickly after first 5 lags. It suggests $q = 5$, $q = 40$ or any value smaller than 40 appropriate for MA(q) component. For the PACF plot, partial correlation dies out quickly after 5 lags. It would be appropriate to select $p = 0$ or $p = 5$ for AR(p) component.

Given the long-range significance in the ACF, I may be dealing with a non-stationary series that needs differencing. Also this plot with a flip to negative autocorrelations and a cyclical pattern that declines over time is a characteristic sign of seasonality in my data. However, STL or classical decomposition fails because time series has no or less than 2 periods, indicating the non-existence of seasonality. So the repeated short-term cycle in 5 lags could be explained by the partial correlation.

```{r, error=TRUE}
start_year <- min(df$year)
end_year <- max(df$year)
frequency <- 1  
ts_cpue <- ts(df$log.CPUE, start=start_year, end=end_year, frequency=frequency)

# Apply STL decomposition
ts_components <- stl(ts_cpue, s.window = "periodic")
plot(ts_components)
```

#### Stationary Check
ARMA(p,q) is performed only if time series data is stationary. So, I will do the stationarity check by ADF.

```{r}
# Kwiatkowski-Phillips-Schmidt-Shin (KPSS) test
kpss_result <- kpss.test(ts_cpue, null = "Level")
print(kpss_result)
```

From ADF test, p-value is estimated as 0.01, smaller than 0.05. We have enough evidence to reject the null hypothesis that the time series is non-stationary.

KPSS with p-value = 0.01 suggests that the log CPUE times series data is non-stationary. This means I can't fit the data with ARMA(p,q) without differencing to achieve stationarity.

#### Differencing
```{r}
# Differencing the time series once
ts_diff <- diff(ts_cpue, differences = 1)
print(kpss.test(ts_diff,null = "Level"))
```

KPSS with p-value = 0.1 suggests that the first-order differenced times series data is stationary. This means I can fit this differenced data with ARMA(p,q).

# Methods

## ARMA
From ADF test, p-value is estimated as 0.01, smaller than 0.05. We have enough evidence to reject the null hypothesis that the time series is non-stationary. KPSS also support the same conclusion that this times series data is stationary. Thus, we could conclude that CPUE on log scale is stationary. This means I can fit the data with ARMA(p,q) without the need to difference or other ways of data transformation to achieve stationarity.

```{r}

aic_table <- function(data,P,Q){
  table <- matrix(NA,(P+1),(Q+1))
  for(p in 0:P) {
    for(q in 0:Q) {
      table[p+1,q+1] <- arima(data,order=c(p,0,q))$aic
    }
  }
  dimnames(table) <- list(paste("AR",0:P, sep=""),
    paste("MA",0:Q,sep=""))
  table
}

bird_aic_table <- aic_table(ts_diff,5,10)
require(knitr)
kable(bird_aic_table,digits=2)

```

The smaller AIC, the better the model to fit the data. Thus, ARMA(0,5) with the lowest AIC value of 204.48 is the best model.

```{r}
aic_table <- function(data,P,Q){
  table <- matrix(NA,(P+1),(Q+1))
  for(p in 0:P) {
    for(q in 0:Q) {
      table[p+1,q+1] <- arima(data,order=c(p,1,q))$aic
    }
  }
  dimnames(table) <- list(paste("AR",0:P, sep=""),
    paste("MA",0:Q,sep=""))
  table
}

bird_aic_table <- aic_table(ts_cpue,5,10)
require(knitr)
kable(bird_aic_table,digits=2)
```

ARIMA(0,1,5) achieves the same result for original log CPUE data.

ARMA(0,5) Diagnosis
```{r}
fit1 <- arima(ts_cpue, order=c(0,0,5))
residuals <- residuals(fit1)

# ACF check
acf(residuals, main="ACF of Residuals")

# Normality check with Q-Q plot
qqnorm(residuals)
qqline(residuals, col="red")

# Shapiro-Wilk test
shapiro.test(residuals)

# Root Check
ma_coefficients <- coef(fit1)[grepl("ma", names(coef(fit1)))]
MA_roots <- polyroot(c(1, ma_coefficients))
#invertible <- all(Mod(MA_roots) > 1)
#print(paste("Is the model invertible? ", invertible))

autoplot(fit1) +
  ggtitle("ARMA(0, 5)")
```

For the ACF plot of the residuals, all autocorrelations are within the confidence bounds (indicated by blue dashed lines), which suggests that there is no significant autocorrelation present. Additionally, the QQ plot shows that the data points roughly align with the reference line (colored red), supporting the assumption of normality. This is further confirmed by a Shapiro-Wilk test for normality, which returns a p-value of 0.62, indicating that the residuals can be considered normally distributed. Moreover, the roots of the characteristic polynomial are outside the unit circle, confirming that the MA(5) process is invertible—a critical condition for model adequacy. These diagnostics collectively verify that the ARIMA(0,1,5) model is well-specified and adheres to the necessary assumptions for valid time series analysis.

## POMP 3

Y(t): bird count, $y_t = NB(mean = \rho B_t, \sigma)$, $B_t$ is bird population size at time t
X(t): unobserved state process, red foxes, Bird, Peak Rodent Year, temperature, precipitation, NAO index

Reference: Lotka-Volterra predator-prey model for the state process: $$X_{t+dt} = \beta X_t - \mu X_t$$

natural born - number of birds eaten by foxes: 
$$B_{t+dt} = \alpha B_t - \beta(R_t) B_{t}F_{t}$$

where, $$\beta(R_t) = \beta if R(t)=0$$ or 
$$ \beta(R_t) = \gamma \beta, \gamma > 1$$ , rodent number affect the rate of fox eating birds, 
$\beta(R_t)$ can be re-written as $$\beta(R_t) =  \gamma^R \beta$$, where R = 0 or 1

Growth Rate - Death Rate of Foxes: 
$$ F_{t+dt} = \delta(t) F_t B^*_t - \theta(t) F_t$$

where, $$\delta(t) = \delta_1 if R(t) = 0, or, \phi \delta_1$$ , $\phi<1$ and whether it is in rodent peak year affect the number of birds and the number of foxes.

$\delta(t)$ can be re-written as $$\delta(t) = \phi^R \delta_1$$, where R = 0 or 1

$\theta(t)$ could include the temperature, NAO about how does it affect the death rate of Foxes.

first, let dt=1 year
then, dt = less than 1 year, maybe 1 month, bird keeps as 100, foxes=10, 15, 17, 15 for each dt until reaching 1 year. 
then include bird in X(t), lecture 16

another way: bird is not in X(t)

$$B_{t+dt} = \alpha B^*_{t+dt} - \beta(R_t) B_{t} F_{t}$$

just do the fox equation, no bird equation

```{r}
# compute the average temperature and preciptation. 
data_frame <- df[,c(1,3,6,7,8:17)]
data_frame$Temp <- rowMeans(data_frame[,c(5:10)], na.rm = TRUE) 
data_frame$Precip <- rowMeans(data_frame[,c(11:14)], na.rm = TRUE) 
data_frame <- data_frame[,c(1,2,3,4,15,16)]
names(data_frame) <- c('year', 'CUPE', 'R', 'NAO', 'Temp', 'Precip')
data_frame$R <- ifelse(data_frame$R == "yes", 1, 0)
data_frame <- transform(data_frame, year = as.numeric(year))  # Ensure 'year' is numeric

```

Euler Approximation of the predator population at time $t+ \delta t$...

$$F_{t+\delta t} = F_t + \delta(t) F_t B_t \delta t - \theta F_t \delta t$$
$$F_{t+\delta t} = F_t + (\phi^R \delta_1 F_t B_t - \theta F_t) \delta t$$

$\delta(t) F_t B_t \delta t$ is the expected number of new predators arising from the Poisson process with rate $\delta(t) F_t B_t$;
$- \theta F_t \delta t$ is the deterministic decrease in predators due to natural death over the time step $\delta t$

$$B_{t+ \delta t} = B_t + \alpha B_t \delta t - \beta(R_t) B_t F_t\delta t$$
$$B_{t+ \delta t} = B_t + (\alpha B_t - \gamma^R \beta B_t F_t) \delta t$$

```{r, warning=FALSE}
meas <- dplyr::select(data_frame, year, reports=CUPE, R)

sir_step<- Csnippet("
  double rateBin, rateBout, rateFin, rateFout;

  rateBin = alpha * B * dt;
  rateBout = pow(gamma, R) * beta_1 * B * Fox * dt;
  rateFin = pow(psi, R) * delta_1 * Fox * B * dt;
  rateFout = theta * Fox * dt;

  // double dF = rbinom(Fox, 1 - exp(-rateF));
  double dFin = rbinom(Fox, 1 - exp(-rateFin));
  double dFout = rbinom(Fox, 1 - exp(-rateFout));

  double dBin = rbinom(B, 1 - exp(-rateBin));
  double dBout = rbinom(B, 1 - exp(-rateBout));

  Fox += dFin - dFout;  
  B += dBin - dBout;
  B = fmax(B, 0.0); 
  Fox = fmax(Fox, 0.0);
  // Rprintf(\"B = %f\\n\\n \", B);

")

# dmeasure: Define the measurement model
sir_dmeas <- Csnippet("
  if (give_log) {
    if (B == 0 & reports != 0) {
      lik = -100000;
    } else {
      // Rprintf(\"rho * B = %f\\n \", rho * B);
      // Rprintf(\"reports = %f\\n\\n \", reports);
      lik = dnbinom_mu(reports, k, rho * B, give_log);
    }
  } else {
    lik = dnbinom_mu(reports, k, rho * B, give_log);
  }
")

# rmeasure: 
sir_rmeas <- Csnippet("
  reports = rnbinom_mu(k, rho * B);
")
## size =  number of successes until the experiment is stopped; success rate / hunting rate:

# initialize:
sir_rinit <- Csnippet("
  Fox = nearbyint(F_0);
  B = nearbyint(B_0);
")

# Define parameters and initial states
params <- c(alpha = 0.5, beta_1 = 0.0001, gamma = 1, delta_1 = 0.01, psi = 3, theta = 0.1, rho = 0.01, k = 20, F_0 = 90, B_0 = 1000)

my_covars <- covariate_table(meas |> dplyr::select(R, year) |> as.data.frame(), times = 'year', order = 'constant')

# create a pomp object
pomp(data = meas |> dplyr::select(-R) |> as.data.frame(),
     rprocess=euler(sir_step,delta.t=1/12), times='year', t0=1871,
    rinit=sir_rinit,
    rmeasure=sir_rmeas,
    dmeasure=sir_dmeas,
    statenames=c("Fox","B"),
    paramnames=c("alpha", "gamma", "delta_1", 'beta_1', "psi", "theta", "rho" ,'k', 'F_0', 'B_0'),
    covar = my_covars,
    partrans = parameter_trans(log = c('alpha', 'gamma', 'delta_1', 'beta_1', 'psi', 'theta', 'rho', 'k', 'F_0', 'B_0'))
  ) -> measSIR

measSIR |>
  pfilter(Np=5, params = params) -> pf

mif2(
  measSIR, 
  Nmif = 10, 
  Np = 2000,
  cooling.fraction.50 = 0.5, 
  rw.sd = rw_sd(
    alpha = 0.02, beta_1 = 0.02, gamma = 0.00, delta_1 = 0.02, psi = 0.02, rho = 0.02, k = 0.02, F_0 = ivp(0.1), B_0 = ivp(0.1), theta = 0.02
  ),
  params = params
) -> mif_out

```

```{r}
statenames <- c('logF', 'logB')
paramnames <- c('a', 'b', 'c', 'gamma', 'alpha', 'Beta', 'sigmaF', 'sigmaB',
                'logF_0', 'logB_0', 'logRho', 'sigma_obs')
obs_names <- c('logCPUE')
# logF_0 = 0 normalizing to assume unit initial population
# sigmaF roughly the noise in a year
param_guess <- c(
  logF_0 = 0,
  logB_0 = 0,
  a = 1,
  b = 2,
  c = 1,
  alpha = 1,
  Beta = 1,
  gamma = 0.5,
  sigmaF = 0.1,
  sigmaB = 0.1,
  logRho = 3,
  sigma_obs = 0.1
)
Csnippet('
  double dwB, dwF;
  dwF = rgammawn(sigmaF, dt);
  dwB = rgammawn(sigmaB, dt);
  logF += (b*R + c*exp(logB)*(1-gamma*R)-a)*dwF;
  logB += (alpha - Beta*exp(logF)*(1-gamma*R))*dwB;
') -> rproc
Csnippet('
  logF = logF_0;
  logB = logB_0;
') -> rinit
Csnippet('
  log.CPUE = rnorm(logB + logrho, signma_obs);
') -> rmeas
Csnippet('
  lik = dnorm(logCPUE, logB + logrho, sigma_obs, give_log);
') -> dmeas

my_covars <- covariate_table(meas |> dplyr::select(R, year) |> as.data.frame(), times = 'year', order = 'constant')

pomp(data = data_frame |> dplyr::select(-R),rprocess=euler(rproc,delta.t=1/12),
     times = 'year', t0 = 1871,
    rinit=rinit,
    # rmeasure=rmeas,
    # dmeasure=dmeas,
    statenames=statenames,
    paramnames=paramnames,
    covar = my_covars
  ) -> measSIR

```